CRUSE to Safe Cycling in Ireland

### 2023-09-21 paper.qmd
Add content from GOT

# # Uncomment for draft versions:
# format: 
#   html: default
#   docx: default

@@

<!-- \[Gary, Suzanne, others input here please.\] -->

@@

The methods used to generate the evidence presented in the CRUSE tool for Ireland build on the Propensity to Cycle Tool (PCT), which was originally funded by the UK's Department for Transport and developed by a multi-university team.
An important feature of the PCT is that it is open source and publicly available (at [www.pct.bike](https://www.pct.bike/)), allowing its use by all stakeholders in the transport planning process [@lovelace2017].
The PCT approach has had major policy impacts, as outlined in Research Excellence Framework (REF) impact case studies, which demonstrate that the tool "revolutionised strategic cycle planning in England and Wales"[^2] by overcoming the barriers to cycling investment imposed by lack of evidence on cycling potential[^3]
.

[^2]: See REF Impact Case Study "Cycle network policy, planning and investment transformed by the Propensity to Cycle Tool" at [results2021.ref.ac.uk](https://results2021.ref.ac.uk/impact/847d1191-7f25-46ba-a399-b481125edc8f) submitted by the University of Leeds.

[^3]: See the REF Impact Case Study "Creating Step Changes in Cycling Policy and Infrastructure Planning across the UK" at [results2021.ref.ac.uk](https://results2021.ref.ac.uk/impact/4BBF3436-FD10-4C75-9791-F5E98AB4411B) submitted by the University of Westminster.

@@

The PCT approach has been applied in other countries, including Ireland (the topic of this paper), Scotland, and Portugal.
In Portugal the 'biclaR' project, based on methods underlying the PCT, has been developed and deployed for the Lisbon metro region.
The resulting evidence is available in an interactive web application hosted at [biclar.tmlmobilidade.pt](https://biclar.tmlmobilidade.pt) [@felix2023].
biclaR includes estimates of impacts, using the World Health Organisation (WHO) 'HEAT for Cycling' tool[^4] and an 'intermodality' scenario that combines cycling with currently available public transit options based on General Transit Feed Specification (GTFS) data.

[^4]: See the Health economic assessment tool for walking and cycling at [www.heatwalkingcycling.org](https://www.heatwalkingcycling.org/#homepage).

>>>>

# Uncomment for draft versions:
format: 
  html: default
  docx: default

@@

A proactive approach to the development of cycling policy and
infrastructure is taking place in Ireland.
The main framework
underpinning these efforts is the Climate Action Plan[^2] and The National
Development Plan[^3]
. Strategic alignment is also outlined by the
Phoenix Park Transport and Mobility Options Study, \"83% of respondents
supported enhanced walking and cycling facilities\"
. The recently
published Greater Dublin Area Transport Strategy builds on this,
suggesting that \"nearly a quarter of adults cycle at least once a week
in the Dublin Metropolitan area\" and that \"cycling in the Dublin area
takes up to 60,000 cars off the road today\"
. (Q for RL: This could be
extrapolated out across other counties in method section if needed?!)
.
Clear evidence of the importance of cycling is provided by the National
Strategic Objective (NSO) from the National Development Plan as \"under
the NSO, sustainable mobility, the Government has committed to investing
€8.6bn in public transport between 2018 and 2027
. This investment will
deliver a wide range of improved infrastructure across both public
transport and active travel, including

[^2]: <https://www.gov.ie/en/publication/7bd8c-climate-action-plan-2023/>

[^3]: <https://www.gov.ie/en/publication/774e2-national-development-plan-2021-2030/>

BusConnects (including
improved cycling infrastructure)\".
This highlights that cycling
infrastructure improvements are to take place concurrently with the
BusConnects project, an entire redesign of the bus network in Dublin and
Cork.
To support this modal shift, TII commissioned AECOM to develop
the CRUSE tool to use as an instrument to understand different cycling
scenarios to de-risk investment in mobility and active travel
infrastructure and promote the benefits of a modal shift towards
cycling.
A key consideration for TII was to ensure both urban and rural
Ireland were targeted for cycling infrastructure in the future.
To this
end, TII extensively reviewed activity on Irish Roads and how the road
network was being used by commuters, cyclists, heavy goods vehicles and
other road users across the Irish road network.
These efforts to analyse
the wider picture of the Irish road network and to identify the
greatest benefits from investment in cycling infrastructure within the
national, regional and local frameworks all provided further
justification to build, design and maintain the CRUSE tool.

The
ability to validate cycling trips, prioritise locations for development
of cycling infrastructure and estimate demand of various cycling
scenarios in Ireland was prioritized by TII given the commitment to
invest in cycling infrastructure by the Irish government.
Furthermore,
TII and Local Authorities would use data generated from the CRUSE tool
to monitor cycling friendliness, comfort and safety in Ireland to ensure
strategic alignment with national, regional and local policies.

When
developing the CRUSE tool, TII, the Institute of Transport Studies
(ITS) Leeds and AECOM to develop the CRUSE Tool to focus on Ireland
specifically(https://cruse.bike/).
ITS Leeds were part of the team that
developed the PCT for England and Wales, with the aim of providing \"an
evidence base to inform cycling investment\".
This experience ensured
that ITS Leeds had extensive experience in developing cutting edge
cycling tools.
The project recognized the importance of an online tool,
using a recognised open-source methodology that will be discussed in
more detail in this paper.
The decision to ensure the tool is opensource
that is publicly available ensures transport planners and members of
the public can utilize the tool for scenario planning to meet their
requirements.
Other use cases in the Irish context includes monitoring
the need for existing cycling infrastructure upgrades, undertaking
appraisal, as well as safety and exposure information in line with the
European Cycling Federation, Dutch Cycling Embassy recommendations and
the European RISM Directive, for reporting collision rates of vulnerable
road users by 2024.

@@

The methods used to generate the evidence presented in the CRUSE tool for Ireland build on the Propensity to Cycle Tool (PCT), which was originally funded by the UK's Department for Transport and developed by a multi-university team.
An important feature of the PCT is that it is open source and publicly available (at [www.pct.bike](https://www.pct.bike/)), allowing its use by all stakeholders in the transport planning process [@lovelace2017].
The PCT approach has had major policy impacts, as outlined in Research Excellence Framework (REF) impact case studies, which demonstrate that the tool "revolutionised strategic cycle planning in England and Wales"[^4] by overcoming the barriers to cycling investment imposed by lack of evidence on cycling potential[^5]
.

[^4]: See REF Impact Case Study "Cycle network policy, planning and investment transformed by the Propensity to Cycle Tool" at [results2021.ref.ac.uk](https://results2021.ref.ac.uk/impact/847d1191-7f25-46ba-a399-b481125edc8f) submitted by the University of Leeds.

[^5]: See the REF Impact Case Study "Creating Step Changes in Cycling Policy and Infrastructure Planning across the UK" at [results2021.ref.ac.uk](https://results2021.ref.ac.uk/impact/4BBF3436-FD10-4C75-9791-F5E98AB4411B) submitted by the University of Westminster.

@@

The PCT approach has been applied in other countries, including Ireland (the topic of this paper), Scotland, and Portugal.
In Portugal the 'biclaR' project, based on methods underlying the PCT, has been developed and deployed for the Lisbon metro region.
The resulting evidence is available in an interactive web application hosted at [biclar.tmlmobilidade.pt](https://biclar.tmlmobilidade.pt) [@felix2023].
biclaR includes estimates of impacts, using the World Health Organisation (WHO) 'HEAT for Cycling' tool[^6] and an 'intermodality' scenario that combines cycling with currently available public transit options based on General Transit Feed Specification (GTFS) data.

[^6]: See the Health economic assessment tool for walking and cycling at [www.heatwalkingcycling.org](https://www.heatwalkingcycling.org/#homepage).
